### 2020-03-03 Hands-on Machine Learning/16_Interpretable-Machine-Learning.rmd
iml
iml

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition_h2o <- as.h2o(attrition)

@@

```{r, echo = T}
ames <- AmesHousing::make_ames()
ames_h2o <- as.h2o(ames)

@@

```{r}
# ensure consistent categorical levels
blueprint <- recipe(Sale_Price ~., data = ames_train) %>%
  step_other(all_nominal(), threshold = 0.005)

@@

### Stacking Overview

Stacking is the process of combining multiple "base" learners (RF, GBM, GLM, etc.) into a "Super Learner" that uses the individual learners to make the final prediction.

The Super Learner Algorithm

1.) Setup the Ensemble

  Specify a list of L base learners (with a specific set of model parameters)

  Specify a meta learning algorithm. This can be any one of the algorithms discussed in the previous chapters, but most often is some form of regularized regression.

2.) Train the ensemble

  Train each of the L base learners on the training set.

  Perform k-fold CV on each of the base learners and collect the cross-validated predictions from each (the same k-fold must be used for each base learner). These predicted values represent $p_1, ..., p_l$.

  The N cross-validated predicted values from each of the L algorithms can be combined to form a new N X L feature matrix (Z)

  Train the meta learning algorithm on level-one data (y = f(Z)). The "ensemble model" consista of the L base learning models and the meta learning model, which can then be used to generate predictions on new data.

3.) Predict on new data.

  To generate ensemble predictions, first generate predictions from the base learners.

  Feed those predictions into the meta learner to generate the ensemble prediction.

### Stacking Existing Models

Restrictions:

1.) All models must be trained on the same training set

2.) All models must be trained with the same number of CV folds.

3.) All models must use the same fold assignment to ensure the same observations are used.

4.) The cross-validated predictions from all of the models must be preserved by setting keep_cross_validation_predictions = T.

```{r}
# Train & cross-validate a GLM model
best_glm <- h2o.glm(
  x = X, y = Y, training_frame = train_h2o, alpha = 0.1,
  remove_collinear_columns = TRUE, nfolds = 10, fold_assignment = "Modulo",
  keep_cross_validation_predictions = TRUE, seed = 123
)

# Train & cross-validate a RF model
best_rf <- h2o.randomForest(
  x = X, y = Y, training_frame = train_h2o, ntrees = 1000, mtries = 20,
  max_depth = 30, min_rows = 1, sample_rate = 0.8, nfolds = 10,
  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
  seed = 123, stopping_rounds = 50, stopping_metric = "RMSE",
  stopping_tolerance = 0
)

# Train & cross-validate a GBM model
best_gbm <- h2o.gbm(
  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.01,
  max_depth = 7, min_rows = 5, sample_rate = 0.8, nfolds = 10,
  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,es

  seed = 123, stopping_rounds = 50, stopping_metric = "RMSE",
  stopping_tolerance = 0
)

# Train & cross-validate an XGBoost model
best_xgb <- h2o.xgboost(
  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.05,
  max_depth = 3, min_rows = 3, sample_rate = 0.8, categorical_encoding = "Enum",
  nfolds = 10, fold_assignment = "Modulo", 
  keep_cross_validation_predictions = TRUE, seed = 123, stopping_rounds = 50,
  stopping_metric = "RMSE", stopping_tolerance = 0
)
```

Train the ensemble

```{r}
# Train a stacked tree ensemble
ensemble_tree <- h2o.stackedEnsemble(
  x = X, y = Y, training_frame = train_h2o, model_id = "my_tree_ensemble",
  base_models = list(best_glm, best_rf, best_gbm, best_xgb),
  metalearner_algorithm = "drf"
)
```

```{r}
get_rmse <- function(model) {
  results <- h2o.performance(model, newdata = test_h2o)
  results@metrics$RMSE
}

list(best_glm, best_rf, best_glm, best_xgb) %>%
  purrr::map_dbl(get_rmse)
```

Stacked Results

```{r}
h2o.performance(ensemble_tree, newdata = test_h2o)@metrics$RMSE
```

```{r}
data.frame(
  GLM_pred = as.vector(h2o.getFrame(best_glm@model$cross_validation_holdout_predictions_frame_id$name)),
  RF_pred = as.vector(h2o.getFrame(best_rf@model$cross_validation_holdout_predictions_frame_id$name)),
  GBM_pred = as.vector(h2o.getFrame(best_gbm@model$cross_validation_holdout_predictions_frame_id$name)),
  XGB_pred = as.vector(h2o.getFrame(best_xgb@model$cross_validation_holdout_predictions_frame_id$name))
) %>% cor()
```

### Stacking a Search Grid

```{r}
# Define GBM hyperparameter grid
hyper_grid <- list(
  max_depth = c(1, 3, 5),
  min_rows = c(1, 5, 10),
  learn_rate = c(0.01, 0.05, 0.1),
  learn_rate_annealing = c(0.99, 1),
  sample_rate = c(0.5, 0.75, 1),
  col_sample_rate = c(0.8, 0.9, 1)
)

# Define random grid search criteria
search_criteria <- list(
  strategy = "RandomDiscrete",
  max_models = 25
)

# Build random grid search 
random_grid <- h2o.grid(
  algorithm = "gbm", grid_id = "gbm_grid", x = X, y = Y,
  training_frame = train_h2o, hyper_params = hyper_grid,
  search_criteria = search_criteria, ntrees = 5000, stopping_metric = "RMSE",
  stopping_rounds = 10, stopping_tolerance = 0, nfolds = 10, 
  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
  seed = 123
)
```

```{r}
# Sort results by RMSE
h2o.getGrid(
  grid_id = "gbm_grid", 
  sort_by = "rmse"
)
```

```{r}
# Grab the model_id for the top model, chosen by validation error
best_model_id <- random_grid@model_ids[[1]]
best_model <- h2o.getModel(best_model_id)
h2o.performance(best_model, newdata = test_h2o)
```

```{r}
# Train a stacked ensemble using the GBM grid
ensemble <- h2o.stackedEnsemble(
  x = X, y = Y, training_frame = train_h2o, model_id = "ensemble_gbm_grid",
  base_models = random_grid@model_ids, metalearner_algorithm = "gbm"
)

# Eval ensemble performance on a test set
h2o.performance(ensemble, newdata = test_h2o)
```

### Auto ML

```{r}
# Use AutoML to find a list of candidate models (i.e., leaderboard)
auto_ml <- h2o.automl(
  x = X, y = Y, training_frame = train_h2o, nfolds = 5, 
  max_runtime_secs = 60 * 120, max_models = 50,
  keep_cross_validation_predictions = TRUE, sort_metric = "RMSE", seed = 123,
  stopping_rounds = 50, stopping_metric = "RMSE", stopping_tolerance = 0
)
```

```{r}
# Assess the leader board; the following truncates the results to show the top 
# 25 models. You can get the top model with auto_ml@leader
auto_ml@leaderboard %>% 
  as.data.frame() %>%
  dplyr::select(model_id, rmse) %>%
  dplyr::slice(1:25)
```

#### Clean-up

>>>>

```{r, attrition_data}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition_h2o <- as.h2o(attrition)

@@

```{r, ames_data}

ames <- AmesHousing::make_ames()
ames_h2o <- as.h2o(ames)

@@

```{r, data_recipe}

# ensure consistent categorical levels
blueprint <- recipe(Sale_Price ~., data = ames_train) %>%
  step_other(all_nominal(), threshold = 0.005)

@@

### h2o ML setup

```{r, train_h2o_models}

# Train & cross-validate a GLM model
best_glm <- h2o.glm(
  x = X, y = Y, training_frame = train_h2o, alpha = 0.1,
  remove_collinear_columns = TRUE, nfolds = 10, fold_assignment = "Modulo",
  keep_cross_validation_predictions = TRUE, seed = 123
)

# Train & cross-validate a RF model
best_rf <- h2o.randomForest(
  x = X, y = Y, training_frame = train_h2o, ntrees = 1000, mtries = 20,
  max_depth = 30, min_rows = 1, sample_rate = 0.8, nfolds = 10,
  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
  seed = 123, stopping_rounds = 5, stopping_metric = "RMSE",
  stopping_tolerance = 0
)

# Train & cross-validate a GBM model
best_gbm <- h2o.gbm(
  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.01,
  max_depth = 7, min_rows = 5, sample_rate = 0.8, nfolds = 10,
  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
  seed = 123, stopping_rounds = 5, stopping_metric = "RMSE",
  stopping_tolerance = 0
)

# Train & cross-validate an XGBoost model

# not avaliable on windows, yet..(3/2/20)

#best_xgb <- h2o.xgboost(
#  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.05,
#  max_depth = 3, min_rows = 3, sample_rate = 0.8, categorical_encoding = "Enum",
#  nfolds = 10, fold_assignment = "Modulo", 
#  keep_cross_validation_predictions = TRUE, seed = 123, stopping_rounds = 50,
#  stopping_metric = "RMSE", stopping_tolerance = 0
#)

# Train a stacked tree ensemble
ensemble_tree <- h2o.stackedEnsemble(
  x = X, y = Y, training_frame = train_h2o, model_id = "my_tree_ensemble_01",
  base_models = list(best_glm, best_rf, best_gbm),
  metalearner_algorithm = "drf"
)
```

### Local Interpretation

Predictions using ensemble tree:

```{r, local_predictions}

# Compute predictions
predictions <- predict(ensemble_tree, train_h2o) %>% as.vector()

# Print the highest and lowest predicted sales price
paste("Observation", which.max(predictions), 
      "has a predicted sale price of", scales::dollar(max(predictions)))

paste("Observation", which.min(predictions), 
      "has a predicted sale price of", scales::dollar(min(predictions)))

# Grab feature values for observations with min/max predicted sales price
high_ob <- as.data.frame(train_h2o)[which.max(predictions), ] %>% select(-Sale_Price)
low_ob  <- as.data.frame(train_h2o)[which.min(predictions), ] %>% select(-Sale_Price)
```

### Model specific vs. model agnostic

```{r, model_features}

# 1) create a data frame with just the features
features <- as.data.frame(train_h2o) %>% select(-Sale_Price)

# 2) Create a vector with the actual responses
response <- as.data.frame(train_h2o) %>% pull(Sale_Price)

# 3) Create custom predict function that returns the predicted values as a vector
pred <- function(object, newdata)  {
  results <- as.vector(h2o.predict(object, as.h2o(newdata)))
  return(results)
}

# Example of prediction output
pred(ensemble_tree, features) %>% head()
```

```{r, iml_dalex_components}
# iml model agnostic object
components_iml <- Predictor$new(
  model = ensemble_tree, 
  data = features, 
  y = response, 
  predict.fun = pred
)

# DALEX model agnostic object
components_dalex <- DALEX::explain(
  model = ensemble_tree,
  data = features,
  y = response,
  predict_function = pred
)
```

### Permutation based feature importance

Implementation

```{r, vip_plot}

vip(
  ensemble_tree,
  train = as.data.frame(train_h2o),
  method = "permute",
  target = "Sale_Price",
  metric = "RMSE",
  nsim = 5,
  sample_frac = 0.5,
  pred_wrapper = pred
)

Partial dependence

```{r, pdp_partial}

# Compute partial dependence values

pd_values <- partial(
  ensemble_tree,
  train = as.data.frame(train_h2o), 
  pred.var = "Gr_Liv_Area",
  pred.fun = function(object, newdata) {
      results <- mean(as.vector(h2o.predict(object, as.h2o(newdata))))
      return(results)
    },
  grid.resolution = 20
)

head(pd_values)  # take a peak

# Partial dependence plot
autoplot(pd_values, rug = TRUE, train = as.data.frame(train_h2o))
```

### Individual conditional expectation

```{r, ice_curves}

# Construct ICE curves
ice_non_centered <- partial(
  ensemble_tree,
  train = as.data.frame(train_h2o), 
  pred.var = "Gr_Liv_Area",
  pred.fun = pred,
  grid.resolution = 20
) %>%
  autoplot(alpha = 0.05, center = FALSE) +
  ggtitle("A) Non-centered ICE curves")

# Construct c-ICE curves
ice_centered <- partial(
  ensemble_tree,
  train = as.data.frame(train_h2o), 
  pred.var = "Gr_Liv_Area",
  pred.fun = pred,
  grid.resolution = 20
) %>%
  autoplot(alpha = 0.05, center = TRUE) +
  ggtitle("B) Centered ICE curves")

# Display plots side by side
gridExtra::grid.arrange(ice_non_centered, ice_centered, ncol = 2)
```

Implementation

### Feature interactions

```{r, pdp_ice_curves}
# Construct c-ICE curves
partial(
  ensemble_tree,
  train = as.data.frame(train_h2o), 
  pred.var = "Gr_Liv_Area",
  pred.fun = pred,
  grid.resolution = 20,
  plot = TRUE,
  center = TRUE,
  plot.engine = "ggplot2"
)
```

```{r, interactions}

interact <- Interaction$new(components_iml)

interact$results %>%
  arrange(desc(.interaction)) %>%
  head()

plot(interact)
```

### Local interpretable model-agnostic explanations

```{r, lime_explainer}

# Create explainer object
components_lime <- lime(
  x = features,
  model = ensemble_tree, 
  n_bins = 10
)

class(components_lime)

# Use LIME to explain previously defined instances: high_ob and low_ob
lime_explanation <- lime::explain(
  x = rbind(high_ob, low_ob), 
  explainer = components_lime, 
  n_permutations = 10,
  dist_fun = "gower",
  kernel_width = 0.25,
  n_features = 10, 
  feature_select = "highest_weights"
)

plot_features(lime_explanation, ncol = 1)
```

### Tuning

```{r, tuning}
# Tune the LIME algorithm a bit
lime_explanation2 <- explain(
  x = rbind(high_ob, low_ob), 
  explainer = components_lime, 
  n_permutations = 5000,
  dist_fun = "euclidean",
  kernel_width = 0.75,
  n_features = 10, 
  feature_select = "lasso_path"
)

# Plot the results
plot_features(lime_explanation2, ncol = 1)

# Two-way PDP using iml
interaction_pdp <- Partial$new(
  components_iml, 
  c("First_Flr_SF", "Overall_Qual"), 
  ice = FALSE, 
  grid.size = 20
)

labels <- interaction_pdp$results %>% filter(First_Flr_SF == max(First_Flr_SF))
plot(interaction_pdp) + 
  ggrepel::geom_label_repel(
    data = labels, 
    aes(label = Overall_Qual),
    label.size = .05, 
    label.padding = .15
)

# Construct c-ICE curves
partial(
  ensemble_tree,
  train = as.data.frame(train_h2o), 
  pred.var = "Gr_Liv_Area",
  pred.fun = pred,
  grid.resolution = 20,
  plot = TRUE,
  center = TRUE,
  plot.engine = "ggplot2"
)
```

### Shapley values

```{r}
# Compute (approximate) Shapley values
(shapley <- Shapley$new(components_iml, x.interest = high_ob, sample.size = 1000))

# Plot results
plot(shapley)

# Reuse existing object
shapley$explain(x.interest = low_ob)

# Plot results
shapley$results %>%
  top_n(25, wt = abs(phi)) %>%
  ggplot(aes(phi, reorder(feature.value, phi), color = phi > 0)) +
  geom_point(show.legend = FALSE)
```

#### Clean up

### 2020-03-03 Mathematics and Statistics for FRM/02_Probabilities.rmd
iml
iml

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Basics}
   \rfoot{\color{headergrey}Chapter 1}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematics and Statistics for Financial Risk Management}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

@@

Solve for $y$, where:

a.) $y = ln(e^5)$

5

b.) $y = ln(1/e)$

$y = ln(1/e) = ln(1) - 1 = 0 - 1 = -1$

c.) $y = ln(10e)$

$y = ln(10e) = ln(10) + ln(e) = ln(10) + 1$

```{r}
log(10) + 1
```

@@

The nominal monthly rate for a loan is quoted at 5%. What is the equivalent annual rate? Semiannual rate? Continious rate?

```{r}
effective_rate <- function(rate, periods) {
   ((1 + rate/periods)^periods - 1)
}
```

Annual:

```{r}
effective_rate(.05, 12) * 100
```

Semiannual:

```{r}
effective_rate(.05, 2) * 100
```

Continous:

```{r}
log(1 + effective_rate(.05, 12)) * 100
```

@@

Over the course of a year, the log return on a stock market index is 11.2%. The starting value of the index is 100. What is the value at the end of the year?

```{r}
100 * exp(.112)
```

@@

You have a portfolio of 10 bonds. In how many different ways can exactly two bonds default? Assume the order in which the bonds default is unimportant.

```{r}
choose(10, 2)
```

@@

What is the present value of a perpetuity that pays $100 per year? Use an actual discount rate of 4%, and assume the first payment will be made in exactly one year.

```{r}
100 * (1/0.04)
```

@@

ABC stock will pay a $1 dividend in one year. Assume the dividend will continue to be paid annually forever and the dividend payments will increase in size at a rate of 5%. Value this stream of dividends using a 6% annual discount rate.

```{r}
1 / (0.06 - 0.05)
```

@@

What is the present value of a 10-year bond with a $100 face value, which pays a 6% coupon annually? Use an 8% discount rate.

```{r}

present_value <- function(face, int.rate, periods, discount.rate) {
   delta <- 1/(1 + discount.rate)

   coupon <- face * int.rate

   discount <- (delta - delta^(periods+1)) / (1 - delta)

   coupon * discount + ( face / (1 + discount.rate)^periods )
}

present_value(100, .06, 10, .08)
```

@@

Solve for $x$, where $e^{e^x} = 10$

```{r}
log(log(10))
```

@@

Calculate the value of the following summation: $\sum^{9}_{i=0}(-0.5)^i$

```{r}
r <- -0.5; n <- 9

(1 - r^n) / (1 - r)
```

@@

The risk department of your firm has 10 analysts. You need to select four analysts to serve on a special audit committee. How many possible groupings of four analysts can be put together?

```{r}
choose(10, 4)
```

long-hand:

```{r}
n <- 10; k <- 4

factorial(10) / ( factorial(k) * factorial(n - k))
```

@@

What is the present value of a newly issued 10-year bond with a notial value of $100 and a 2% annual coupon? Assume a constant 5% annual discount rate and no risk of default.

```{r}
present_value(100, .02, 10, 0.05)
```

>>>>

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Probabilities}
   \rfoot{\color{headergrey}Chapter 2}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematics and Statistics for Financial Risk Management}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@
